\subsection{Details on Random Problem Generation}\label{appB:SAT}

As we alluded to in Section~\ref{sec:sat},
some issues may arise if the restrictions on the randomly generated formula are too strict:
\begin{itemize}
    \item
        The sample space might be empty or very sparse.
        In practice, it seems to the user as if the generator got stuck,
        usually resulting in the process being killed by the user. 
        For example,
        consider the restriction on polarities of propositional
        variables. % (Sect.~\ref{sec:sat}).
        Combined with the other restrictions,
        it is impossible to get a formula that contains atomic propositions
        of purely positive and purely negative polarity at the same time.

    \item
        The second issue manifests less drastically but is perhaps more problematic:
        the sample space may be too uniform,
        leading to the generation of trivial and/or very similar
        formulas. 
        In particular, we encountered this problem
        when we restricted the number of models to exactly one, or zero.
        We note that there simply are not that many ways to rule out 8 interpretations
        using only 7 connectives.
\end{itemize}





Regarding the filtering of generated formulas using the constraints discussed in Section~\ref{sec:satfo},
we did not require very efficient algorithms since
the formulas under consideration are very small.
For example, for the restriction on the number of models we used a naive satisfiability test
based on evaluating the formula under each possible interpretation.
An advantage of this is that the addition of new filters/constraints is relatively easy.




For the random problem generation setting of Section~\ref{sec:satfo},
we used design principles from the Haskell library
QuickCheck~\cite{ClaessenHughes:2000:QuickCheck}.
%which we used for the first prototype.
However, because of our many filtering criteria, we wanted the generator to support backtracking.
To this end,
we created a simple typeclass \texttt{MonadChoose} in the style of the monad transformer library (mtl),
with a single primitive operation \texttt{choose} for choosing an element from a list of possible choices:
\lstset{
    language=Haskell,
    basicstyle=\small\ttfamily,
    deletekeywords={MonadPlus,filter,length},
    columns=fixed,
}
\begin{lstlisting}
class MonadPlus m => MonadChoose m where
  -- Choose element with uniform probability
  choose :: [a] -> m a
\end{lstlisting}

Our generator implementations are generic over the monad, constrained by \texttt{MonadChoose}.
The following listing shows (a slightly simplified) part of the inference generator
discussed in Section~\ref{sec:fo}.
\begin{lstlisting}
genExamInference :: MonadChoose m => m Inference
genExamInference = do
  -- Define signature (partially omitted)
  let vars = ["x", "y", "z"]
  let opts = GenOptions{ vars = vars, ... }

  -- Choose variables to appear in l1 and l2
  v1 <- choose vars
  v2 <- choose (filter (/= v1) vars)

  -- Generate literals
  -- l1: exactly one occurrence of v1
  l1 <- mfilter ((==1) . length . toListOf variables)
        $ genUninterpretedLiteral opts{ vars = [v1] }
  -- l2: at least two occurrences of v2
  l2 <- mfilter ((>=2) . length . toListOf variables)
        $ genEqualityLiteral opts{ vars = [v2] }
  -- l3: ground literal
  l3 <- genUninterpretedLiteral opts{ vars = [] }

  -- (rest omitted)
  return inference

genEqualityLiteral, genUninterpretedLiteral
  :: MonadChoose m => GenOptions -> m Literal
-- (literal generators omitted)
\end{lstlisting}

We used two concrete implementations to evaluate generators:
\begin{enumerate}
    \item
        \texttt{RandomChoice}, a monad that implements \texttt{choose}
        as uniform random choice with backtracking support.
        Conceptually, this is like the standard list monad
        where \texttt{choose} works like the regular monadic bind for lists
        except that it shuffles the list with a random permutation first.
        This evaluation method is used to generate random exams.
    \item
        The standard list monad to enumerate the sample space.
        This second evaluation method helps verifying that the sample space is sufficiently large.
\end{enumerate}

\subsection{Weights and Precedences for Ground Superposition}\label{appB:FO}

The weights and precedences used to generate the KBOs for the superposition problem
from Section~\ref{sec:gs}
are displayed in Table~\ref{tab:ground-sup}.
The upper part of the table shows all weight and precedence combinations,
denoted as $w_{i, I}, p_{i, I}$ for $i \in \{1, 2, 3, 4\}$
and $I \in \{f, g\}$ (for convenience, the table contains both $w_{i, f}$
and $w_{i, g}$, as well as $p_{i, f}$ and $p_{i, g}$ for all values of~$i$).
The lower part of the table displays the values of $i_1, I_1; i_2, I_2$; and $i_3, I_3$,
corresponding to the three weight and precedence combinations selected for
each instance of the clause set.\todo{Make this easier to read?}

\begin{table}
\begin{center}
\begin{tabular}{r@{\hskip 0.5em}c c c c@{\hskip 0.5em} |@{\hskip 0.5em} l@{\hskip 1em} ||@{\hskip 0.5em} r@{\hskip 0.5em} c c c c@{\hskip 0.5em} |@{\hskip 0.5em} l}
  weight of: & $f$ & $g$ & $a$ & $b$ & precedence
  & weight of: & $f$ & $g$ & $a$ & $b$ & precedence \\ \hline
  $w_{1,f}:$  & 1   & 3   & 2   & 1   & $p_{1,f}: a \gg b \gg f \gg g$ &
  $w_{1,g}:$  & 3   & 1   & 2   & 1   & $p_{1,g}: a \gg b \gg g \gg f$ \\ \hline
  $w_{2,f}:$  & 0   & 3   & 2   & 1   & $p_{2,f}: f \gg a \gg g \gg b$ &
  $w_{2,g}:$  & 3   & 0   & 2   & 1   & $p_{2,g}: g \gg a \gg f \gg b$ \\ \hline
  $w_{3,f}:$  & 0   & 1   & 3   & 1   & $p_{3,f}: f \gg a \gg b \gg g$ &
  $w_{3,g}:$  & 1   & 0   & 3   & 1   & $p_{3,g}: g \gg a \gg b \gg f$ \\ \hline
  $w_{4,f}:$  & 1   & 2   & 3   & 1   & $p_{4,f}: g \gg f \gg a \gg b$ &
  $w_{4,g}:$  & 2   & 1   & 3   & 1   & $p_{4,g}: f \gg g \gg a \gg b$ \\ \hline
  % original table:
  %$w_1(h_1) = w_1(b) = 1, w_1(a) = 2, w_1(h_2) = 3$ & \quad $p_1: a \gg b \gg h_1 \gg h_2$ \\
  %$w_2(h_1) = 0, w_2(b) = 1, w_2(a) = 2, w_2(h_2) = 3$ & \quad $p_2: h_1 \gg a \gg h_2 \gg b$ \\
  %$w_3(h_1) = 0, w_3(b) = w_3(h_2) = 1, w_3(a) = 3$ & \quad $p_3: h_1 \gg a \gg b \gg h_2$ \\
  %$w_4(b) = w_4(h_1) = 1, w_4(h_2) = 2, w_4(a) = 3$ & \quad $p_4: h_2 \gg h_1 \gg a \gg b$ \\
\end{tabular}

\hspace*{0.5em}

\begin{tabular}{l@{\hskip 1.05em} || c | c | c}
\hline
condition & $i_1, I_1$ & $i_2, I_2$ & $i_3, I_3$ \\ \hline
$F \not = G$ and $X \not= Y$ & $1,E$ & $2, E$ & $3, E$ \\
$F \not = G$ and $X = Y$ & $1, H$ & $2, E$ & $4, H$ \\
$F = G$ and $X \not= Y$ & $1, H$ & $2, H$ & $3, E$
\end{tabular}
\hspace*{0.5em}
\caption{Weights and precedences for the ground superposition problem.}
\label{tab:ground-sup}
\end{center}
\end{table}
