% \begin{example}\label{ex:sat}
\subsubsection*{Problem 1 (SAT).}
    Consider the formula:
    \[
        % Random number generator seed: Just 123
        % Index: 36
        % Satisfiable? True
        % Valid? False
        % Polarities: fromList [(p,Both),(q,Both),(r,Pos)]
        % Models: [{¬p,q,r}, {¬p,¬q,r}, {¬p,¬q,¬r}]
        ( r \land \lnot ( q \rightarrow p ) ) \lor ( q \leftrightarrow \lnot ( p \rightarrow q ) )
    \]
    \begin{enumerate}[(a)]
        \item
            Which atoms are pure in the above formula?
        \item\label{subtask:sat-cnf}
            Compute a clausal normal form $C$ of the above formula by
            applying the CNF transformation algorithm with naming and
            optimization based on polarities of subformulas.
        \item
            Decide the satisfiability of the computed CNF formula $C$
            by applying the DPLL method to $C$. If $C$ is
            satisfiable, give an interpretation which
            satisfies it.
            %    \qed
    \end{enumerate}
% \end{example}

% \begin{example}\label{ex:fo}
\subsubsection*{Problem 2 (Non-Ground Superposition and Redundancy).}
    Consider the following inference:
    \begin{center}
        \begin{tabular}{c}
            $P ( h ( g ( g ( d , d ) , b ) ) ) \lor  \lnot P ( h ( f ( d ) ) ) \lor f ( d ) \neq h ( g ( a , a ) )$ \\
            $\lnot P ( h ( g ( x , b ) ) ) \lor f ( d ) \neq h ( g ( y , y ) )$ \\
            \hline
            $\lnot P ( h ( f ( d ) ) ) \lor f ( d ) \neq h ( g ( a , a ) )$ \\
        \end{tabular}
    \end{center}
    % \[
    %     % Random number generator seed: Just 123
    %     % Index: 36
    %     % Substitution: {x -> g(d,d),y -> a}
    %     \prftree{P ( h ( g ( g ( d , d ) , b ) ) ) \lor  \lnot P ( h ( f ( d ) ) ) \lor f ( d ) \neq h ( g ( a , a ) )}{ \lnot P ( h ( g ( x , b ) ) ) \lor f ( d ) \neq h ( g ( y , y ) )}{ \lnot P ( h ( f ( d ) ) ) \lor f ( d ) \neq h ( g ( a , a ) )}
    % \]
    in the non-ground superposition inference system $\textrm{Sup}$
    (including the rules of the non-ground binary resolution inference system $\textrm{BR}$),
    where $P$ is a predicate symbol, $f$, $g$, $h$ are function symbols, $a$, $b$, $d$ are constants, and $x$, $y$ are variables.
    %
    \begin{enumerate}[(a)]
        \item
            Prove that the above inference is a sound inference of $\textrm{Sup}$.
        \item
            Is the above inference a simplifying inference of $\textrm{Sup}$?
            Justify your answer based on conditions of clauses being redundant.
    \end{enumerate}
% \end{example}

% \begin{example}
\subsubsection*{Problem 3 (SMT).}
    Consider the formula:
\begin{align*}
%
& %
% ['=', 'b', 'c']
b = c
%
\land
%
% ['not', ['=', ['f', ['+', 'b', '1']], ['+', 'b', '2']]]
f(b + 1)  \neq  b + 2
%
\land
%
% ['=', ['select', 'A', ['f', ['+', 'c', '1']]], 'c']
\mathit{read}(A,f(c + 1)) = c
~\\~
& %
%
\land
%
% ['or', ['=', ['select', 'A', ['f', ['+', 'b', '1']]], ['+', 'b', '3']], ['=', ['select', ['store', 'A', ['+', 'b', '2'], ['f', 'c']], ['f', ['+', 'c', '1']]], ['+', 'c', '2']]]
(\mathit{read}(A,f(b + 1)) = b + 3
\\ & \hphantom{~\land (} \lor \mathit{read}(\mathit{write}(A,b + 2,f(c)),f(c + 1)) = c + 2)
%
\end{align*}
    where
    $b$, $c$ are constants,
    $f$ is a unary function symbol,
    $A$ is an array constant,
    $\mathit{read}$, $\mathit{write}$ are interpreted in the array theory,
    and
    $+$, $1$, $2$, $3$, $\ldots$ are interpreted in the standard way over the integers.
    \medskip

    \noindent
    Use the Nelson-Oppen decision procedure in conjunction with DPLL-based
    reasoning in the combination of the theories of arrays,
    uninterpreted functions, and linear integer arithmetic.
    Use the decision procedures for the theory of arrays and the theory of uninterpreted functions and use
    simple mathematical reasoning for deriving new equalities among the
    constants in the theory of linear integer arithmetic.
    If the formula is satisfiable, give an interpretation that satisfies the formula.\bigskip
% \end{example}

% \begin{example}
\subsubsection*{Problem 4 (Ground Superposition).}
Consider the KBO ordering $\succ$ generated
by the precedence $f \gg a \gg b \gg g$
and the weight function $w$ with $w(f) = 0, w(b) = 1, w(g) = 1, w(a) = 3$.
Let $\sigma$ be a well-behaved selection function wrt $\succ$.
Consider the set $S$ of ground formulas:
\begin{align*}
%
% ['or', ['=', ['f', ['g', 'b']], 'a'], ['=', ['f', ['g', 'a']], 'a']]
& f(g(b)) = a \lor f(g(a)) = a
~\\~
% ['=', ['g', 'b'], 'a']
& g(b) = a
~\\~
% ['=', ['g', 'a'], 'a']
& g(a) = a
~\\~
% ['or', ['not', ['=', ['g', 'b'], ['g', 'b']]], ['not', ['=', ['f', 'a'], 'a']]]
& g(b)  \neq  g(b) \lor f(a)  \neq  a
%
\end{align*}
\noindent Show that $S$ is unsatisfiable by applying saturation on $S$ using an
inference process based on the ground superposition
calculus $\textrm{Sup}_{\succ,\sigma}$ (with the inference rules of
binary  resolution $\textrm{BR}_{\sigma}$ included).
 Give details on what literals
 are selected and which terms are maximal. \bigskip
% \end{example}
